---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# namecount

<!-- badges: start -->
<!-- badges: end -->

The **namecount** package contains 5 functions:**getNamesCount**, **getNamesStart**, **getNamesUnique**, **countStart**, and **countNamesUnique**. These functions can be used to count the number of names in a specified column or retrieve names from a specified column.The goal of the **getNamesCount** package is to provide a quick and easy way to filter your to better understand your data set without having to write out a long
function yourself.

## Installation

You can install the development version of namecount from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stat545ubc-2023/namecount")
```

## getNamesCount - Example 1

This is a basic example of the `getNamesCount` function from the **namecount** package using a basic data set: 

```{r example}
# calling the 'namecount' package
library(namecount)
# basic sample data frame containing numeric and categorical columns 
dfSample <- data.frame(
fruit = c("apple", "banana", "cherry", "coconut", "dragonfruit"),
numFruit = c(17, 90, 35, 49, 102))
# using the 'getNamesCount' function 
answer1.0 <- getNamesCount(dfSample, fruit, numFruit, 50)

print(answer1.0)
```
## getNamesCount - Example 2

Looking at the summary statistics from the **penguins** data set from the *palmerpenguins* package you can see that the mean bill length collected in the data is 43.92mm. 
```{r palmerpenguins}
sumP <- summary(palmerpenguins::penguins)

print(sumP)
```
Now I would like to know how many observations there are in the data set for species of penguins with a bill length below the mean collected. This is where I would use the **namecount** function:
```{r}
library(namecount)

answer2.0 <- getNamesCount(palmerpenguins::penguins, species, bill_length_mm, 43.92)

print(answer2.0)
```
The **getNamesCount** function produced a number, this number is the total species observations that have a corresponding bill length measurement below the *numericalValue* that was set at the mean (43.92mm) 

### Conclusion

If you would like to see more examples of usage for the **namecount** package please refer to the package vignette or function help pages. 

## Website

[Link to pkgdown website](https://stat545ubc-2023.github.io/namecount/)

## References 

The source of the **penguins** data set from the `palmerpenguins` R package used here: 

Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. doi:10.5281/zenodo.3960218, R package version 0.1.0, https://allisonhorst.github.io/palmerpenguins/.