---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# namecount

<!-- badges: start -->
<!-- badges: end -->

The goal of the **getNamesCount** function from the **namecount** package is to provide a quick and easy way to count the number of observations in a categorical column of your data set that are less than a given numeric value of a numeric column. 

This is a quick and easy way to better understand the types of observations in your data set without having to write out a long function yourself. 

## Installation

You can install the development version of namecount from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stat545ubc-2023/namecount")
```

## Example 1

This is a basic example using a basic data set: 

```{r example}
# calling the 'namecount' package
library(namecount)
# basic sample data frame containing numeric and categorical columns 
dfSample <- data.frame(
fruit = c("apple", "banana", "cherry", "coconut", "dragonfruit"),
numFruit = c(17, 90, 35, 49, 102))
# using the 'getNamesCount' function 
answer1.0 <- getNamesCount(dfSample, fruit, numFruit, 50)

print(answer1.0)
```
## Example 2

Looking at the summary statistics from the **penguins** data set from the *palmerpenguins* package you can see that the mean bill length collected in the data is 43.92mm. 
```{r palmerpenguins}
sumP <- summary(palmerpenguins::penguins)

print(sumP)
```
Now I would like to know how many observations there are in the data set for species of penguins with a bill length below the mean collected. This is where I would use the **namecount** function:
```{r}
library(namecount)
answer2.0 <- getNamesCount(palmerpenguins::penguins, species, bill_length_mm, 43.92)

print(answer2.0)
```
The **getNamesCount** function produced a number, this number is the total species observations that have a corresponding bill length measurement below the *numericalValue* that was set at the mean (43.92mm) 
