---
title: "namecountVingette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{namecountVingette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `namecount` package provides functionality to count names and filter names that meet specified criteria within a dataset. This vignette introduces the main functions and provides example usage.


## Installation

You can install the development version of namecount from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stat545ubc-2023/namecount")
```

## Package Functions 

The **namecount** package includes the following functions:

1. `countNamesUnique`: Count the number of unique values in a categorical column.
2. `countStart`: Count the number of categorical observations starting with a specific letter.
3.  `getNamesCount`: Count names that fall within a numerical value range.
4.  `getNamesStart`: Find unique variables in a column starting with a specific letter.
5.  `getNamesUnique`: Display all unique names within a categorical column.

## Examples

### countNamesUnique Function

The **countNamesUnique** function can be used to count the number of unique names found in the categorical column in the data set 

```{r}
library(namecount)
library(palmerpenguins)

# Count unique species in the penguins dataset

countSpecies <- countNamesUnique(palmerpenguins::penguins, "species")

print(countSpecies)
```

### countStart Function

The **countStart** function can be used to count the number of observations in a categorical column that start with a specified letter.

```{r}
# For the example here we will try to count names that start with the letter 'g'

dfSample <- data.frame(
fruit = c("apple", "banana", "cherry", "coconut", "dragonfruit"),
numFruit = c(17, 90, 35, 49, 102))

cNameCount <- countStart(dfSample, "fruit", "g")

print(cNameCount)
```


### getNamesCount Function

The **getNamesCount** can be used to count names from a specified categorical colunm that fall within a numerical value range.

```{r}
# Count penguins with bill length less than 300 mm

countBills <- getNamesCount(palmerpenguins::penguins, species, bill_length_mm, 35)

print(countBills)
```

### getNamesUnique Function

The **getNamesUnique** function can be used to identify the unique observation names in a specified column. 

```{r}
# Display unique island names from the penguins dataset

uniqueIslands <- getNamesUnique(palmerpenguins::penguins, "island")

print(uniqueIslands)
```

### getNamesStart Function

The **getNamesStart** function can be used to retrieve the names in a specified column that start with a given letter.

```{r}
# Get the names of observations from the island colunm that start with the letter 'T'

tNames <- getNamesStart(palmerpenguins::penguins, "island", "T")

print(tNames)
```

## Conclusion

The **namecount** package provides a set of functions for working with categorical data, making it easier to count, filter, and explore unique names within datasets.

If you are looking for more details and additional functionalities for the functions in the **namecount** package, refer to the package documentation and function help pages.

## References 

The source of the **penguins** data set from the `palmerpenguins` R package used here: 

Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. doi:10.5281/zenodo.3960218, R package version 0.1.0, https://allisonhorst.github.io/palmerpenguins/.
